---
title: "crg_examples"
author: "beyza"
date: "2025-05-20"
output: ggplot_exercises
---


```{r}
#install.packages('BiocManager')
#BiocManager::install('multtest')
#install.packages('Seurat')
#install.packages("ggplot2")
#install.packages("devEMF")

```

ggplot exercises
1. dowloading data
```{r}

acs <- readRDS(url("https://sscc.wisc.edu/sscc/pubs/dvr/acs.rds"))
acs_small <- readRDS(url("https://sscc.wisc.edu/sscc/pubs/dvr/acs_small.rds"))
library(ggplot2)
```

2. ggplot Building Blocks
aes (short for aesthetic)
| Term                  | Meaning                                     |
| --------------------- | ------------------------------------------- |
| `ggplot()`            | Starts a new plot                           |
| `aes()`               | Connects variables to the plot              |
| `geom_*()`            | Chooses the type of plot (bar, point, etc.) |
| `+`                   | Adds parts to the plot                      |
| `inherit.aes = FALSE` | Ignores the general mapping settings        |


```{r}
ggplot(acs, aes(x = edu)) +
  geom_bar()
```

geom_bar() draws a bar chart and counts the values for you automatically.
"stat_count() counts the values first, then shows them as bars.
```{r}
ggplot(acs, aes(x = edu)) +
  stat_count()
```
```{r}
ggplot(acs, aes(x = age, y = log(income))) +
  geom_abline(color = "red", intercept = 7, slope = .05, size = 3, linetype = 3) +
  geom_point() +
  geom_hline(color = "gold", yintercept = 10, size = 3, linetype = 2) +
  geom_smooth(se = F, size = 3)
```
```{r}
ggplot(acs, aes(x = edu)) +
  geom_bar()
```

3.1.  one variable
filter(!is.na(edu)) --> for filterşng na
|> ---->pipe oparoru for doing code step by step
coord_flip() ----> for changing x and y axis
```{r}
acs |> 
  filter(!is.na(edu)) |> 
  ggplot(aes(x = edu)) +
  geom_bar() +
  coord_flip()

```

mine
```{r}
edu_without_na <- na.omit(acs)
ggplot(edu_without_na, aes(x= edu))+ 
geom_bar() +
coord_flip()
```

3.2. Continuous
3.2.1 Histograms
```{r}
ggplot(acs, aes(x = age)) +
  geom_histogram()
```

binwith for frequence
```{r}
ggplot(acs, aes(x = age)) +
  geom_histogram(binwidth = 10)
```


3.2.2 Density Plots
```{r}
ggplot(acs, aes(x = age)) +
  geom_density()
```


```{r}
ggplot(acs, aes(x = age)) +
  geom_density(adjust = .2)
```

4. Two Variables
```{r}
library(ggplot2)
library(dplyr)
library(scales)
```

4.1 One Discrete, One Continuous
4.1.1 Barplots


```{r}
acs |> 
  filter(!is.na(edu)) |> 
  group_by(edu) |> 
  summarize(income = mean(income, na.rm = T)) |> 
  ggplot(aes(x = edu, y = income)) +
  geom_col() + 
  coord_flip() +
  geom_text(aes(label = round(income)), 
            hjust = -.2)
```


4.1.2 Boxplots
coord_flip() can be used to rotate the plot 90 degrees.
```{r}
acs |> 
  filter(!is.na(edu), !is.na(income)) |> 
  ggplot(aes(x = edu, y = income)) +
  geom_boxplot() +
  coord_flip()
```

The boxplot has covered up the violin plot, so we can reduce the width of the boxplot with the width argument.
```{r}
ggplot(acs, aes(x = race, y = age)) +
  geom_violin() +
  geom_boxplot(width = .25) +
  coord_flip()
```

4.2 Two Continuous
4.2.1 Scatterplots
```{r}
ggplot(acs, aes(x = age, y = income)) +
  geom_point()
```

```{r}
ggplot(acs, aes(x = age, y = income)) +
  geom_point(size = .5)
```
```{r}
ggplot(acs, aes(x = age, y = income)) +
  geom_point(alpha = .1)
```
We can also add a linear fit of y ~ x by using geom_smooth() and setting its method to "lm".
```{r}
ggplot(acs, aes(x = age, y = income)) +
  geom_point(alpha = .1) +
  geom_smooth(method = "lm")
```


4.3 Two Discrete
4.3.1 Scatterplots
```{r}
table(acs$edu, acs$race, useNA = "ifany")
```
```{r}
prop.table(table(acs$edu, acs$race, useNA = "ifany"))
```

```{r}
ggplot(acs, aes(x = race, y = edu)) +
  geom_jitter()
```

4.3.2 Barplots
Setting the position argument of geom_bar() to "dodge" places the bars side by side.
```{r}
ggplot(acs, aes(x = race, fill = edu)) +
  geom_bar(position = "dodge")
```


```{r}
ggplot(acs, aes(x = race, fill = edu)) +
  geom_bar(position = "stack")
```

using facet_grid(). Faceting allows us to split our one plot into several panels
 use round() within our ggplot call to round perc to the nearest tenth.
```{r}
acs |> 
  filter(!is.na(edu)) |> 
  group_by(race, edu) |> 
  summarize(n = n()) |> 
  mutate(perc = 100*n/sum(n)) |> 
  ggplot(aes(x = race, y = perc, fill = edu)) +
  geom_col() +
  facet_grid(~ edu) + 
  coord_flip() +
  ylim(c(0, 90)) +
  geom_text(aes(label = round(perc, 1)), hjust = -.1)
```

5. Three Variables
More than two variables can be visualized without resorting to 3D plots by mapping the third variable to some other aesthetic, or by creating a separate plot (“facet”) for each of its values.

5.1 With Aesthetics
Note that since female is numeric, ggplot created a legend with a continuous color scale.
```{r}
ggplot(acs_small, aes(x = age, y = income, color = female)) +
  geom_point()
```
To correct this, we can either change our dataframe to make female a character or factor vector, or we can temporarily specify it as such when we create our plot.
```{r}
ggplot(acs_small, aes(x = age, y = income, color = as.factor(female))) +
  geom_point()
```


```{r}
ggplot(acs_small, aes(x = age, y = income, color = edu)) +
  geom_point()
```


We can supply facet_wrap() with the formula ~ edu. 
```{r}
ggplot(acs_small, aes(x = age, y = income)) +
  geom_point() +
  facet_wrap(~ edu)
```

6.1. Adding and Changing Labels
Labels can be added to a plot with labs() by assigning a character string to the aesthetic (x, y, shape, etc.).
```{r}
ggplot(acs_small, aes(x = age, y = income, 
                       shape = as.factor(female))) +
  geom_point() +
  labs(title = "Income by Age, Sex, and Education",
       caption = "Random sample of\nACS Dataset",
       x = "Age",
       y = "Income (in US Dollars)",
       shape = "Sex") +
  facet_wrap(~ edu) +
  scale_shape_discrete(labels = c("Male", "Female"))
```

6.2. Changing Text and Axes
Text size, font, angle, and position are all customizable. We can tell theme() what we would like to chang
```{r}
ggplot(acs_small, aes(x = age, y = income, 
                       shape = as.factor(female))) +
  geom_point() +
  labs(title = "Income by Age, Sex, and Education",
       caption = "Random sample of\nACS Dataset",
       x = "Age",
       y = "Income (in US Dollars)",
       shape = "Sex") +
  facet_wrap(~ edu) +
  scale_shape_discrete(labels = c("Male", "Female")) +
  theme(axis.title.x = element_text(size = 10, face = "italic"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.text.y = element_text(angle = 20),
        axis.ticks.length.x = unit(1, "cm"),
        axis.ticks.x = element_line(arrow = arrow()),
        axis.ticks.y = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(angle = 15, vjust = -.2),
        strip.text = element_blank()) +
  scale_x_continuous(breaks = seq(0, 100, 20))
```

6.3 Changing Themes
| Theme             | Features                                       |
| ----------------- | ---------------------------------------------- |
| `theme_gray()`    | Gray background, default theme                 |
| `theme_bw()`      | White background, simple and professional look |
| `theme_minimal()` | Very simple design with few lines              |
```{r}
myTheme <- theme(panel.border = element_rect(color = "black", fill = NA),
                 panel.grid = element_line(color = "#EEEEEE"),
                 panel.background = element_rect(fill = NA),
                 legend.key = element_rect(fill = NA))

ggplot(acs_small, aes(x = age, y = income, 
                       shape = as.factor(female))) +
  geom_point() +
  labs(title = "Income by Age and Sex",
       caption = "Random sample of\nACS Dataset",
       x = "Age",
       y = "Income (in US Dollars)",
       shape = "Sex") +
  scale_shape_discrete(labels = c("Male", "Female")) +
  myTheme
```


7. Saving Plots
devEMF is an R package used to save high-quality plots as EMF (vector image) files for documents or presentations.
```{r}
library(devEMF)

ggplot(acs, aes(x = age)) +
  geom_density()

ggsave("plot.emf", width = 6, height = 4, device = {function(filename, ...) devEMF::emf(file = filename, ...)})
ggsave("plot.pdf", width = 6, height = 4)
ggsave("plot.png", width = 6, height = 4)
```


8. Exercises
8.1 Working Backwards
```{r}
ggplot(acs_small, aes(x = age, y = income)) +
  geom_point() +
  geom_smooth(se = F, size = 3)
```

```{r}
ggplot(acs_small, aes(x = age, y = income, color= edu)) +
geom_point()

```


```{r}
ggplot(acs_small, aes(x= age, y= maritalStatus)) +
geom_col()
```























