---
title: "r_example"
author: "beyza"
date: "2025-06-02"
output: r_example
---
Get the length of the “lynx” dataset
```{r}
View(lynx)
```


```{r}
length(lynx)
```
1b. Create a vector of ordered “lynx” index numbers 
```{r}
order(lynx)
```
Get 2 vectors (index positions and absolute values) with all “lynx” observations smaller than 500
```{r}
which(lynx < 500)#for position
```

```{r}
subset(lynx, lynx < 500)#for values
```

2a. Get a Histogram of the “lynx” dataset

2b. Adjust the bin size to 7 bins

2c. Remove the labs and change the bins to 2 alternating colors

2d. Add a subtitle and divide the main title in 2 lines
```{r}
hist(lynx, breaks = 7,
     col= c("red", "blue"),
     sub= "Histogram of r tutorials",
     xlab="",
     ylab= "")
```
Extract Sepal.Length from the “iris” dataset and call the resulting vector mysepal
```{r}

mysepal <- iris$Sepal.Length
```

3b. Get the summation, mean, median, max and min of mysepal
```{r}
sum(mysepal);
mean(mysepal);
median(mysepal);
min(mysepal);
max(mysepal)
summary(mysepal)
```


4a. Install and load library “dplyr”
4b. Call help for function arrange of “dplyr”
4c. Deinstall “dplyr”,
```{r}
#install.packages("dplyr")
library(dplyr)
#remove.packages("dplyr")
```

6a. Get a subset of “mtcars” with transmission type: manual, and call it mysubset
```{r}
mtcars2 <- mtcars$am
mtcars3 <- subset(mtcars, select= am)
```

6b. Extract the first 2 variables of the first 2 observations of mysubset
```{r}
mtcars_2future <- mtcars[c(1.2),c(3,4)]
```

7a. Get the first 9 observations of “mtcars”
```{r}
mtcars_9 <- mtcars[(1:9),]
#head(mtcars, 9)
```

Get the “mtcars” dataset ordered according to the increasing amount of “carb”
hint for 7b:
```{r}
library(dplyr)
mtcars_arrange <- arrange(mtcars, carb)

```


8a. Get the means of the first 2 columns in the “iris” dataset by Species
```{r}
by(iris[1:2], iris$Species, colMeans)
```
8b. Create vector x which is the alternation of 1 and 2, 75 times, check the length
```{r}
x = rep(1:2, times=75); length(x)
```

8c. Attach x to iris as dataframe “irisx”, check the head
```{r}
iris$x <- x
```

8d. Get the colsums of columns 1,3,4 in the “irisx” dataset by the new x variable
```{r}
by(iris[c(1,3,4)], iris$x, colSums)
```

9a. How many observations in “lynx” are smaller than 500?
```{r}
sum(lynx < 500)
```

9b. How many observations in “iris” have a Sepal.Length greater or equal 5?
```{r}
sum(iris$Sepal.Length >=5)
```

10a. Plot a simple xy plot with “iris” Sepal.Length vs. Sepal.Width
10b. Enlarge the scale limits: y from 0 – 5, x from 3 – 9
10c. Add a text of your choosing, in red, at the lower part of the plot
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width, ylim=c(0,5), xlim=c(3,9))
```




